### 2025-01-10 scripts/08.figures_markdown/archive/FigS11_rareVariants.Rmd

---
title: "Figure S11: Role of rare variant enrichment"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Load data & functions
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
supplement_figures_dir <- paste0(base_dir, "outputs/updated_figures/figures/supplement/")
tables_dir <- paste0(base_dir, "outputs/updated_figures/tables/")
#analysis_objects_dir <- paste0(base_dir, "outputs/updated_figures/objects/Fig4_characterizeGRCCA/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WGCNA_res.R"))

## Load DGE data
load(paste0(base_dir, "objects/DE_results.RDS")) # df_de_res (generated in A.DGE_analysis.R)

## Load GRCCA data
load(paste0(base_dir, "objects/GRCCA_results.Rdata")) # df_lvs, df_y_res, df_x_res

## Pull rare variants for plots
rare_variants <- benchmarking_lists[["SCZ (rare)"]]
```

```{r setup_labels}
## Create dataframe of labels for rare variants on plots
df_rare_variant_labs <- df_x_res %>% 
    filter(ensembl_gene_id %in% rare_variants) %>% 
    left_join(df_de_res) %>% 
    mutate(
        p_adj_lab = ifelse(p_adj < 0.01, format(p_adj, digits = 3), round(p_adj, 2)),
        GRCCA_lab = paste0("R=", round(pearsons_r, 2), "; FDR=", p_adj_lab),
        DGE_lab = paste0("L2FC=", round(log2fold_change*10, 2), "; p=", round(pvalue, 2))
    ) %>% 
    dplyr::select(ensembl_gene_id, gene_symbol, GRCCA_lab, DGE_lab)
```

### A | Expression direction of rare variants

```{r FigS11a.prep}
## Set factor levels for rare variant genes using structure correlation
rare_variant_order <- df_x_res %>% 
    filter(ensembl_gene_id %in% rare_variants) %>% 
    arrange(-pearsons_r) %>% 
    pull(gene_symbol)

## Pull normalized and corrected expression values for each rare variant-associated gene
df_figS11a <- df_vsd_regress %>% 
    dplyr::select(sample, any_of(rare_variants)) %>% 
    pivot_longer(contains("ENSG"), names_to = "ensembl_gene_id", values_to = "expr") %>% 
    left_join(df_ensembl_to_symbol, by = join_by(ensembl_gene_id)) %>% 
    left_join(df_rare_variant_labs, by = join_by(ensembl_gene_id, gene_symbol)) %>% 
    mutate(
        dx = case_when(
            str_detect(sample, "control") ~ "Control",
            str_detect(sample, "bipolar") ~ "BD",
            str_detect(sample, "mdd") ~ "MDD",
            str_detect(sample, "schizo") ~ "SCZ"
        ) %>% 
            factor(levels = names(dx_colors)),
        gene_symbol = factor(gene_symbol, levels = rare_variant_order)
    )
```

```{r FigS11a.plot}
figS11a <- df_figS11a %>% 
    ggplot(aes(x = dx, y = expr)) +
    geom_point(aes(color = dx), position = position_jitter(width = 0.2), size = 0.3) +
    geom_boxplot(aes(color = dx, fill = dx), alpha = 0.1, outlier.shape = NA) +
    geom_hline(yintercept = 0, color = "black", lty = 2) +
    geom_text(aes(label = GRCCA_lab), x = Inf, y = -Inf, vjust = -0.5, hjust = 1,
              check_overlap = TRUE, size = 2.75) +
    scale_fill_manual(values = dx_colors, guide = "none") +
    scale_color_manual(values = dx_colors, guide = "none") +
    facet_wrap(vars(gene_symbol), scales = "free_y", nrow = 3) +
    labs(x = NULL, y = "Corrected expression value",
         title = "A | Rare variant-associated gene expression values") +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 0.9),
        strip.text = element_text(size = 9)
    )
```

```{r FigS11a.save, fig.height = 4, fig.width = 5}
figS11a
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(supplement_figures_dir, "S11A.rareVariant_gene_expression_values", .x), width = 5, height = 4)
)
```

### B | Rare variants in GRCCA structure correlation and DGE L2FC distributions

```{r FigS11b.prep}
## Create df of all struct cor & L2FC for each gene
df_figS11b <- df_x_res %>% 
    left_join(df_de_res, by = join_by(ensembl_gene_id, gene_symbol)) %>%

    # pivot DGE and GRCCA stats for facet wrapping
    pivot_longer(cols = c(pearsons_r, log2fold_change), names_to = "analysis", values_to = "value") %>% 
    mutate(analysis = ifelse(analysis == "pearsons_r", "GRCCA structure correlation", "DGE L2FC")) %>%

    # identify significant genes in each analysis
    mutate(significant = case_when(
        analysis == "GRCCA" & p_adj < 0.05 ~ 1,
        analysis == "DGE" & pvalue < 0.05 ~ 1,
        TRUE ~ 0
    ) %>% 
        factor
    ) %>% 
    arrange(significant, abs(value))

## Create separate df that only contains points to plot
df_figS11b_points <- df_figS11b %>% 
    dplyr::filter(ensembl_gene_id %in% rare_variants)

## Plotting specificities
scale_max <- df_x_res %>% pull(pearsons_r) %>% abs %>% max
jitter_pos <- position_jitter(width = 0.1, seed = 123)
```

```{r FigS11b.function}
## Write function to plot distributions
f_plot_gene_points_distribution <- function(df, df_points) {
    df %>%
        ggplot(aes(x = value)) +
        geom_density() +
        geom_vline(xintercept = 0, color = "darkgrey") +
        geom_point(data = df_points, 
                   mapping = aes(x = value, y = -1, fill = value, color = significant), 
                   pos = jitter_pos, shape = 21, size = 2
        ) +
        geom_text_repel(data = df_points,
                        mapping = aes(x = value, y = -1, label = gene_symbol),
                        pos = jitter_pos, 
                        size = 3, box.padding = 0.5, max.overlaps = 50, force = 10, min.segment.length = 0
        ) +
        #facet_wrap(vars(analysis), scales = "free") +

        scale_fill_gradientn(colors = gene_weight_color_scale, guide = "none") +
        scale_color_manual(values = c("gray", "black"), guide = "none") +
        labs(x = NULL, y = NULL, 
             title = NULL) +
        theme(axis.text.y = element_blank(),
              axis.ticks.y = element_blank(),
              axis.line.y = element_blank()
        )
}
```

```{r FigS11b.plot}
f_plot_gene_points_distribution(df = df_figS11b %>% filter(analysis == "DGE L2FC"), 
                                df_points = df_figS11b_points %>% filter(analysis == "DGE L2FC")
) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-0.27, 0.27), guide = "none")

```

```{r FigS11b.save, fig.height = 2, fig.width = 4}
figS11b
```