### 2025-01-10 scripts/08.figures_markdown/archive/TRY-transcript_GRCCA_res.Rmd

---
title: "Potential supplement figures: Transcript-level GRCCA results"
author: "Rachel Smith"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(fig.width = 2.5, fig.height = 2, collapse = TRUE, warning = FALSE, message = FALSE, echo = FALSE)

## Set paths and load data
base_dir <- "~/Documents/PhD/projects/sgacc_wgcna_grcca/"
figures_dir <- paste0(base_dir, "outputs/figures/Fig5_GRCCAtranscriptsRes/")
tables_dir <- paste0(base_dir, "outputs/tables/")
analysis_objects_dir <- paste0(base_dir, "outputs/objects/Fig5_GRCCAtranscriptsRes/")

source(paste0(base_dir, "scripts/full_analysis_scripts/scripts/load_WTCNA_res.R"))

## Load data objects for plots
objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}
```

### Y weights

```{r, fig.height = 3.5, fig.width = 10}
p_covariates <- df_y_res %>% 
    pivot_longer(c(z_score, pearsons_r), names_to = "metric", values_to = "value") %>% 
    mutate(metric = ifelse(metric == "z_score", "weight z-score", "structure correlation")) %>% 
    mutate(significance = case_when(
        metric == "weight z-score" & abs(value) > 2 ~ "*",
        metric == "structure correlation" & p_adj < 0.05 & p_adj > 0.01 ~ "*",
        metric == "structure correlation" & p_adj < 0.01 & p_adj > 0.001 ~ "**",
        metric == "structure correlation" & p_adj < 0.001 ~ "***",
        TRUE ~ ""
    )
    ) %>%

    ggplot(aes(x = value, y = reorder_within(covariate, value, model))) +
    geom_point(aes(size = -log10(p_adj), fill = value), shape = 21) +
    geom_vline(xintercept = 0, lty = 2) +
    geom_text(aes(label = significance), size = 4, vjust = -0.15, color = "black") +
    facet_wrap(model ~ metric, scales = "free", nrow = 1) +

    scale_fill_gradientn(colors = gene_weight_color_scale) +
    scale_size_continuous(range = c(2, 6)) +
    scale_y_reordered() +
    guides(fill = guide_colorbar(title = NULL),
           size = guide_legend(title = "-log10(FDR)")) +
    coord_cartesian(clip = "off") +
    labs(y = NULL, x = NULL, 
         title = "Covariate results") +
    theme(legend.position = "bottom", 
          legend.justification = "center",
          legend.box = "horizontal",
          #legend.justification = "center",
          #legend.key.size = unit(0.2, 'cm')
          legend.key.height = unit(0.20, "cm"),
          legend.key.width = unit(1.0, "cm")
    )
p_covariates
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "covariate_weight_structCor", .x), width = 10, height = 3.5)
)
```

### X weights

```{r}
df_x_res %>% 
    pivot_wider(id_cols = c(ensembl_transcript_id, transcript_symbol), names_from = model, values_from = pearsons_r) %>%

    ggplot(aes(x = includeDim2, y = regressDim2)) +
    geom_point() +
    geom_abline(lty = 2) +
    geom_smooth(method = "lm") +
    stat_cor()
```

### Risk gene GSEA

```{r, fig.height = 3, fig.width = 5}
p_risk_genes <- df_grcca_fgsea_benchmark %>% 
    mutate(class = ifelse(str_detect(benchmark_list, "SCZ"), "SCZ", "other") %>% factor(levels = c("SCZ", "other"))) %>% 
    mutate(benchmark_list = str_remove_all(benchmark_list, "SCZ|\\)|\\(") %>% 
               str_trim %>% 
               str_replace(", ", ",\n")
    ) %>% 
    mutate(label_color = ifelse(padj < 0.05, "white", "black")) %>% 
    mutate(label = round(NES, 3)) %>%

    ggplot(aes(x = benchmark_list, y = 1)) +
    geom_tile(aes(fill = NES)) +
    geom_text(aes(label = label, color = I(label_color)), size = 3) +
    facet_wrap(class ~ model, scales = "free", nrow = 2) +
    scale_fill_gradientn(colors = gene_weight_color_scale, limits = c(-2, 2)) +
    guides(fill = guide_colorbar(title.hjust = 0.5, title.position = "left")) +
    labs(x = NULL, y = NULL,
         title = "Risk gene enrichment") +
    theme(axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          axis.line = element_blank(),
          axis.text.x = element_text(margin = margin(t = -3)),
          legend.margin = margin(r = -15),
          legend.position = "left",
          legend.justification = "top",
          legend.title = element_text(angle = 90),
          legend.key.height = unit(0.75, "cm"),
          legend.key.width = unit(0.15, "cm")
    )
p_risk_genes
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "risk_gene_matrices", .x), width = 5, height = 3)
)
```

### GO GSEA

```{r, fig.height = 8, fig.width = 9}
p_go_gsea <- df_grcca_fgsea_go %>% 
    dplyr::filter(padj < 0.05 & model == "regressDim2") %>% 
    mutate(direction = ifelse(NES < 0, "down", "up")) %>% 
    group_by(direction) %>% 
    mutate(label = ifelse(padj < 0.031, term, "")) %>% 
    #mutate(label = ifelse(term %in% fig4c_labels, str_wrap(term, 25), "")) %>%

    ggplot(aes(x = padj, y = NES)) +
    geom_point(aes(size = size, fill = NES), color = "gray", shape = 21) +
    geom_text_repel(aes(label = label), size = 3, force = 5,
                    box.padding = 0.5, min.segment.length = 0.1, max.overlaps = 30) +
    geom_hline(yintercept = 0, color = "gray") +
    facet_wrap(vars(ontology), nrow = 2) +
    scale_fill_gradientn(colors = rev(brewer.rdbu(100)), limits = c(-3, 3), guide = "none") +
    scale_size_continuous(range = c(2, 6)) +
    scale_x_reverse(breaks = c(0, 0.025, 0.05)) +
    coord_cartesian(clip = "off") +
    guides(size = guide_legend(title = "n genes in path", title.hjust = 0.5, title.position = "top")) +
    labs(x = "Enrichment FDR", y = "NES",
         title = "GRCCA transcript enrichments (regressDim2)") +
    theme(
        legend.position = "bottom",
        legend.key.size = unit(0.3, "cm")
    )
p_go_gsea
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0(figures_dir, "GO_GSEA", .x), width = 9, height = 8)
)
```

### Example transcript plot (relationship with Dim2 & relationship with SCZ)

```{r}
load(paste0(base_dir, "outputs/objects/FigS2_drugMCA/drug_MCA_results.Rdata")) # df_var_loadings, df_ind_loadings, df_eig

## Identify transcript to plot
top_transcript <- df_x_res %>% 
    filter(model == "includeDim2") %>% 
    top_n(10, pearsons_r) %>% 
    pull(ensembl_transcript_id)

## Combine transcript expression with Dim 2 scores & diagnosis
df_dim2_tx <- df_ind_loadings %>% 
    dplyr::select(sample, dim2) %>% 
    left_join(
        df_vsd_regress_filt %>% 
            dplyr::select(sample, all_of(top_transcript))
    ) %>% 
    left_join(
        df_covariates %>% dplyr::select(sample, dx)
    ) %>% 
    pivot_longer(contains("ENST"), names_to = "ensembl_transcript_id", values_to = "expr") %>% 
    left_join(df_transcript_to_gene) %>% 
    distinct() %>% 
    mutate(transcript_symbol = ifelse(is.na(transcript_symbol), ensembl_transcript_id, transcript_symbol))

## Plot correlation between Dim2 score and transcript expression
df_dim2_tx %>% 
    ggplot(aes(x = dim2, y = expr)) +
    geom_point() +
    geom_smooth(method = "lm") +
    stat_cor() +
    facet_wrap(vars(transcript_symbol), scales = "free")

## Plot expression levels across diagnostic groups
df_dim2_tx %>% 
    ggplot(aes(x = dx, y = expr)) +
    geom_point(aes(color = dx), position = position_jitter(width = 0.2), size = 0.5) +
    geom_boxplot(aes(color = dx, fill = dx), alpha = 0.1, outlier.shape = NA) +
    geom_hline(yintercept = 0, color = "black", lty = 2) +
    scale_fill_manual(values = dx_colors, guide = "none") +
    scale_color_manual(values = dx_colors, guide = "none") +
    facet_wrap(vars(transcript_symbol), scales = "free_y") +
    labs(x = NULL, y = "Corrected expression value",
         title = "Transcript expression values") +
    theme(axis.text.x = element_text(angle = 45, hjust = 0.9))

```

```{r}
df_dim2_txAll <- df_ind_loadings %>% 
    dplyr::select(sample, dim2) %>% 
    left_join(df_vsd_regress_filt)
# Run all correlations
df_x_res_dim2cor <- df_x_res %>% 
    mutate(
        dim2_cor = map(
            .x = ensembl_transcript_id,
            .f = ~ cor.test(df_dim2_txAll$dim2, df_dim2_txAll[[.x]])$estimate
        )
    ) %>% 
    unnest(cols = c(dim2_cor))
# Plot
df_x_res_dim2cor %>% 
    ggplot(aes(x = dim2_cor, y = pearsons_r)) +
    geom_point() +
    stat_cor()
```

### Gene & transcript correlation

```{r, fig.height = 3, fig.width = 3.5}
load(paste0(analysis_objects_dir, "../Fig3_GRCCAres/GRCCA_results.Rdata")) # df_x_res
df_x_res_genes <- df_x_res

objects <- list.files(analysis_objects_dir)
objects <- objects[!str_detect(objects, "null|archive")]
for (obj in objects){
  print(paste0("loading ", obj, "...."))
  load(paste0(analysis_objects_dir, obj))
}

## Map transcript to gene and correlate
p <- df_x_res %>% 
    filter(model == "regressDim2" & !is.na(ensembl_gene_id)) %>% 
    group_by(ensembl_gene_id) %>% 
    summarise(pearsons_r = max(pearsons_r)) %>% 
    dplyr::rename("transcript_res" = "pearsons_r") %>%

    left_join(df_x_res_genes %>% 
                  dplyr::select(ensembl_gene_id, gene_symbol, pearsons_r) %>% 
                  dplyr::rename("gene_res" = "pearsons_r")
    ) %>% 
    mutate(
        risk_gene = ifelse(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, prioritized)"]] | ensembl_gene_id %in% benchmarking_lists[["SCZ (rare)"]], "yes", "no"),
        label = ifelse(risk_gene == "yes", gene_symbol, "")
    ) %>% 
    arrange(risk_gene) %>%

    ggplot(aes(x = gene_res, y = transcript_res)) +
    geom_vline(xintercept = 0, color = "gray") +
    geom_hline(yintercept = 0, color = "gray") +
    geom_point(aes(fill = gene_res + transcript_res, color = risk_gene, size = risk_gene), shape = 21) +
    geom_text_repel(aes(label = label), size = 2.5, max.overlaps = 50) +
    stat_cor() +
    geom_smooth(method = "lm", color = "black", linewidth = 0.75) +
    xlim(c(-0.6, 0.6)) +
    ylim(c(-0.6, 0.6)) +
    scale_fill_gradientn(colors = gene_weight_color_scale, guide = "none") +
    scale_color_manual(values = c("yes" = "black", "no" = "transparent")) +
    scale_size_manual(values = c("yes" = 1.5, "no" = 0.5)) +
    labs(x = "Gene GRCCA structure correlations", y = "Transcript GRCCA structure correlations",
         title = "Correspondence between gene & \ntranscript GRCCA results")
map(
  .x = c(".png", ".pdf"),
  .f = ~ ggsave(paste0("~/Downloads/gene_transcript_GRCCA_cor", .x), width = 3.5, height = 3)
)
```

```{r}
## Risk gene transcripts
df_x_res %>% 
    filter(ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]])

## Sig GRCCA transcripts
sig_grcca_genes <- df_x_res_genes %>% 
    filter(abs(z_score) >= 2 & p_adj < 0.05) %>% 
    pull(ensembl_gene_id)
df_x_res %>% 
    filter(ensembl_gene_id %in% sig_grcca_genes & ensembl_gene_id %in% benchmarking_lists[["SCZ (common, broad)"]])
```